---
title: "Stats2Project2"
author: "Aniketh V"
date: "7/22/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load the libraries
```{r}
library(dplyr)
library(ggplot2)
library(tidyr)
library(magrittr)
library(stringr)
library(corrplot)
library(caret)
library(glmnet)
library(Lahman)
library(mice)
library(car)
library(MASS)
library(ROCR)
library(caret)
library(mosaic)
library(ggmosaic)
library(dplyr)
library(ggplot2)
library(tidyr)
library(SDMTools)
library(readr)
library(digest)
library(ISLR)
library(car) 
library(leaps)
library( Matrix)
library(foreach)
library(glmnet)
library(VIM)
library(mice)
library(corrgram)
library(car)
library(tidyverse)
library(limma)
library(gridExtra)
library(MASS)
library(mvtnorm)
library(class)
library(caret)
library(e1071)
```

Loading dataset
```{r}
#imported the main data
Bank_Fix <- read.csv("bank-additional-full.csv",sep=";",header=TRUE, strip.white = TRUE, na.strings = c("unknown"))

#Separted the header string into separate columns
#Bank_Fix = mainBank %>% separate(age.job.marital.education.default.housing.loan.contact.month.day_of_week.duration.campaign.pdays.previous.poutcome.emp.var.rate.cons.price.idx.cons.conf.idx.euribor3m.nr.employed.y, c("age", "job", "marital", "education", "default", "housing", "loan", "contact", "month", "day_of_week", "duration", "campaign", "pdays", "previous", "poutcome", "emp.var.rate", "cons.price.idx", "cons.conf.idx", "euribor3m", "nr.employed", "y"), ";", extra = "merge")

Bank_Fix$age = as.integer(Bank_Fix$age)
Bank_Fix$nr.employed = as.integer(Bank_Fix$nr.employed)
Bank_Fix$euribor3m = as.double(Bank_Fix$euribor3m)
Bank_Fix$cons.conf.idx = as.double(Bank_Fix$cons.conf.idx)
Bank_Fix$cons.price.idx = as.double(Bank_Fix$cons.price.idx)
Bank_Fix$emp.var.rate = as.double(Bank_Fix$emp.var.rate)
Bank_Fix$previous = as.integer(Bank_Fix$previous)
Bank_Fix$pdays = as.integer(Bank_Fix$pdays)
Bank_Fix$campaign = as.integer(Bank_Fix$campaign)
Bank_Fix$duration = as.integer(Bank_Fix$duration)
Bank_Fix$y = as.factor(Bank_Fix$y)
Bank_Fix$poutcome = as.factor(Bank_Fix$poutcome)
Bank_Fix$term_deposit = Bank_Fix$y
Bank_Fix$default = as.factor(Bank_Fix$default)
Bank_Fix$housing = as.factor(Bank_Fix$housing)
Bank_Fix$marital = as.factor(Bank_Fix$marital)
Bank_Fix$loan = as.factor(Bank_Fix$loan)
Bank_Fix$job = as.factor(Bank_Fix$job)


Bank_Fix$y = NULL
#Bank_Fix[Bank_Fix=="unknown"]<-NA


#Bank_Fix$default = str_replace(Bank_Fix$default, "unknown", "NA")
# Bank_Fix$housing = str_replace(Bank_Fix$housing, "unknown", "NA")
# Bank_Fix$marital = str_replace(Bank_Fix$marital, "unknown", "NA")
# Bank_Fix$loan = str_replace(Bank_Fix$loan, "unknown", "NA")

#find out the data types
dplyr::glimpse(Bank_Fix)


#Imputed NA values
tempData <- mice(Bank_Fix,m=1,maxit=0,method ='logreg',seed=500)
Bank_Fix_Imp <- complete(tempData,1)

```

```{r, split}
set.seed(100)
split_percent = .70
trainIndices = sample(1:dim(Bank_Fix_Imp)[1],round(split_percent * dim(Bank_Fix_Imp)[1]))
train = Bank_Fix_Imp[trainIndices,]
test = Bank_Fix_Imp[-trainIndices,]
summary(train)

table(train$term_deposit)

set.seed(1000)
down_train <- downSample(x = train[, -ncol(train)],
                         y = train$term_deposit)
table(down_train$Class)

library(DMwR)

set.seed(9560)
smote_train <- SMOTE(term_deposit ~ ., data  = train)                         
table(smote_train$term_deposit)

##library(ROSE)

#set.seed(9560)
#rose_train <- ROSE(Class ~ ., data  = imbal_train)$data                         
#table(rose_train$Class)
```

```{r}
#Logistic regression on Normal Train set before SMOTE
full.original.log = glm(term_deposit~.,family="binomial",data=train)
step.original.log = full.original.log %>% stepAIC(trace=FALSE)
summary(step.original.log)
exp(cbind("Odds ratio" = coef(step.original.log), confint.default(step.original.log, level = 0.95)))
vif(step.original.log)

#smote Train data  step wise
full.log = glm(term_deposit~.,family="binomial",data=smote_train)
step.log = full.log %>% stepAIC(trace=FALSE)
summary(step.log)
exp(cbind("Odds ratio" = coef(step.log), confint.default(step.log, level = 0.95)))
vif(step.log)

#smote Train data lasso 
dat.train.x <- model.matrix(term_deposit ~ .,smote_train)[,-1]
dat.train.y<-smote_train$term_deposit
cvfit <- cv.glmnet(dat.train.x, dat.train.y, family = "binomial", type.measure = "class", nlambda = 1000)
plot(cvfit)
coef(cvfit, s = "lambda.min")

#final lasso model
finalLassoModel<-glmnet(dat.train.x, dat.train.y, family = "binomial",lambda=cvfit$lambda.min)
cvfit$lambda.min

##performing prediction using both models


dat.test.x<-model.matrix(term_deposit ~ .,test)[,-1]

fit.pred.lasso <- predict(finalLassoModel, newx = dat.test.x, type = "response")
fit.pred.step<-predict(step.log,newdata=test,type="response")

cutoff<-0.6
class.lasso<-factor(ifelse(fit.pred.lasso>cutoff,"Yes","No"),levels=c("No","Yes"))
class.step<-factor(ifelse(fit.pred.step>cutoff,"Yes","No"),levels=c("No","Yes"))


#Confusion Matrix for Lasso
conf.lasso<-table(class.lasso,test$term_deposit)
print("Confusion matrix for LASSO")
## [1] "Confusion matrix for LASSO"
conf.lasso

conf.step<-table(class.step,test$term_deposit)
print("Confusion matrix for Stepwise")
## [1] "Confusion matrix for Stepwise"
conf.step

sum(diag(conf.lasso))/sum(conf.lasso)
sum(diag(conf.step))/sum(conf.step)




results.lasso<-prediction(fit.pred.lasso, test$term_deposit)
roc.lasso = performance(results.lasso, measure = "tpr", x.measure = "fpr")
plot(roc.lasso,colorize = TRUE)
abline(a=0, b= 1)



Bank_Fix_Imp


fit <- lm( term_deposit ~ .^2, data=Bank_Fix_Imp )


#DO later
grid = 10^seq(10,-2, length = 100)
lasso.mod = glmnet(dat.train.x,dat.train.y,alpha = 1, lambda = grid)

cv.out = glmnet(dat.train.x,dat.train.y, alpha = 1)
plot(cv.out)
lambdaFunct = cv.out$lambda.min
lasso.pred = predict(lasso.mod, s = lambdaFunct, newx = dat.train.x)

MSE_lasso = mean((dat.train.y - lasso.pred)^2)
MSE_lasso
coef(lasso.mod, s = lambdaFunct)
```


EDA
```{r}
#Some initial EDAs
summary(Bank_Fix_Imp)

#Split catergorical and continuous data
Bank_Conti <- Bank_Fix_Imp[, !sapply(Bank_Fix_Imp, is.factor)]
Bank_categ <- Bank_Fix_Imp[, sapply(Bank_Fix_Imp, is.factor)]

#Boxplots
boxplot(Bank_Conti$age)
boxplot(Bank_Conti$nr.employed)
boxplot(Bank_Conti[,2,3])

#ScatterPlot
pairs(Bank_Conti[,1:10], pch = 19)

#PCA
pc.result<-prcomp(Bank_Conti,scale=FALSE)
pc.scores<-pc.result$x
cor(pc.scores)
pc.result$rotation
 
par(mfrow=c(1,2))
eigenvals<-(pc.result$sdev)^2
plot(1:10,eigenvals/sum(eigenvals),type="l",main="Scree Plot",ylab="Prop. Var. Explained")
cumulative.prop<-cumsum(eigenvals/sum(eigenvals))
plot(1:10,cumulative.prop,type="l",main="Cumulative proportion",ylim=c(0,1))
par(mfrow=c(1,1))

#Correlation Matrix
# Computing the p value of correlations
cor.mtest <- function(mat, ...) {
    mat <- as.matrix(mat)
    n <- ncol(mat)
    p.mat<- matrix(NA, n, n)
    diag(p.mat) <- 0
    for (i in 1:(n - 1)) {
        for (j in (i + 1):n) {
            tmp <- cor.test(mat[, i], mat[, j], ...)
            p.mat[i, j] <- p.mat[j, i] <- tmp$p.value
        }
    }
  colnames(p.mat) <- rownames(p.mat) <- colnames(mat)
  p.mat
}

correlation = cor.mtest(Bank_Conti[,1:10])
mat = cor(Bank_Conti[, 1:10])

corrplot(correlation, type = "upper", order = "hclust", insig = "label_sig", p.mat = mat, sig.level = 0.05)

p.mat <- cor.mtest(who_imp_conti[,2:20])
correlation <- cor(who_imp_conti[,2:20])
#Correlation plot with significance level of 0.05
corrplot(correlation, type="upper", order="hclust", insig = "label_sig", p.mat = p.mat, sig.level = 0.05)

correlation <- cor(Bank_Conti)
corrplot(correlation)

#Format the data
attach(Bank_Fix_Imp)
x_vars = model.matrix(term_deposit~ ., data = Bank_Fix_Imp)[,-1]
y_vars = Bank_Fix_Imp$term_deposit
grid=10^seq(10,-2, length = 100)

set.seed(100)
train = sample(1:nrow(x_vars), nrow(x_vars)/2)
x_test = (-train)
y_test = y_vars[x_test]

Lasso = glmnet(x, y, alpha = 1, lambda = grid)


## Summary on Job variable, customers job status
summary(Bank_Fix_Imp$job)
catnames = names(Bank_Fix_Imp)[sapply(Bank_Fix_Imp, class) == "factor"]

Bank_Fix_Imp$term_deposit = as.factor(Bank_Fix_Imp$term_deposit)

spineplot(x = Bank_Fix_Imp$job, y = Bank_Fix_Imp$term_deposit, xlab = "Job", ylab = "y",
          main = "Job vs Y", col = c("lightblue", "coral"), xaxlabels = levels(Bank_Fix_Imp$job))
chisq.test(Bank_Fix_Imp$job, Bank_Fix_Imp$term_deposit)
CrossTable(Bank_Fix_Imp$job, Bank_Fix_Imp$term_deposit)

##job is dependent on term_deposit
##marital is dependent on term_deposit

summary(Bank_Fix_Imp$marital)
chisq.test(Bank_Fix_Imp$marital, Bank_Fix_Imp$term_deposit)
CrossTable(Bank_Fix_Imp$marital, Bank_Fix_Imp$term_deposit)
chisq.test(Bank_Fix_Imp$education, Bank_Fix_Imp$term_deposit)
chisq.test(Bank_Fix_Imp$default, Bank_Fix_Imp$term_deposit)
chisq.test(Bank_Fix_Imp$housing, Bank_Fix_Imp$term_deposit)
chisq.test(Bank_Fix_Imp$loan, Bank_Fix_Imp$term_deposit)
chisq.test(Bank_Fix_Imp$contact, Bank_Fix_Imp$term_deposit)
chisq.test(Bank_Fix_Imp$month, Bank_Fix_Imp$term_deposit)
chisq.test(Bank_Fix_Imp$day_of_week, Bank_Fix_Imp$term_deposit)
chisq.test(Bank_Fix_Imp$poutcome, Bank_Fix_Imp$term_deposit)
CrossTable(Bank_Fix_Imp$poutcome, Bank_Fix_Imp$term_deposit)
##marital is dependent on term_deposit
## contact has some difference in "yes" and "no" among its categories (cellular and telephone). cellular with 14.7% and 5.2% for "yes" rsponse
## P-value of  Chi-Square Test suggests that the variable "contact" has a  relationship with response variable. We can keep this variable for final analysis
## Day of the week has some difference in "yes" and "no" among its categories. Most of the calls were on Thursday (12.1%) and other days are close to 10% 
## P-value of  Chi-Square Test suggests that the variable "day_of_week" has a  relationship with response variable. We can keep this variable for final analysis
## 65.1% of customers where previous outcome was "Success" has a response of "yes"  
## 14.2% of customers where previous outcome was "failure" has a response of "yes"  
## 8.8% of customers who were not contacted has a response of "yes"  
## P-value of  Chi-Square Test suggests that the variable "poutcome" has a  relationship with response variable. We can keep this variable for final analysis
##### We need to keep below variables in the predictive model
## job   marital  education  contact  month  day_of_week  poutcome
##### Below variables will not be included in the predictive model as there is no significance with response variable
## default  housing  loan


#multicolliniarity check
bank.model<-lm(age~duration+campaign+pdays+previous+emp.var.rate+cons.price.idx+cons.conf.idx+euribor3m+nr.employed, data=Bank_Fix_Imp)
summary(bank.model)
vif(bank.model)
# removed varaible euribor3m which has VIF 63.51
bank.model1<-lm(age~duration+campaign+pdays+previous+emp.var.rate+cons.price.idx+cons.conf.idx+nr.employed, data=Bank_Fix_Imp)
summary(bank.model1)
vif(bank.model1)
#removed varaible emp.var.rate which has VIF 24.12
bank.model2<-lm(age~duration+campaign+pdays+previous+cons.price.idx+cons.conf.idx+nr.employed, data=Bank_Fix_Imp)
summary(bank.model2)
vif(bank.model2)

#### We can remove variables emp.var.rate and euribor3m as these variables are highly correlated with nr.employed
corrgram(Bank_Fix_Imp, order=TRUE,
         upper.panel=panel.cor, lower.panel=panel.pie, main="Bank data Correlogram")


## euribor3m and nr.employed are highly correlated (0.95)
## emp.var.rate and euribor3m are highly correlated (0.97)
## emp.var.rate and nr.employed are highly correlated (0.91)

summary(Bank_Fix_Imp$age)
Bank_Fix_Imp %>%  ggplot(aes(x = age, fill = y, color=y)) + geom_bar() + ggtitle("Distribution of Age") + xlab("Age") + 
  scale_x_continuous(breaks = seq(0, 100, 5))
ggplot(Bank_Fix_Imp, aes(x = y, y=age, fill=y)) + geom_boxplot() + ggtitle("Distribution of Age") + xlab("Response") + ylab ("Age")
```

## The minimum and maximum values are 17 and 98 and distribution of age is slightly right screwed 
## Highest concentration of values between 22 and 60 and distribution of values between 22 and 60 is normal


```{r }
## Summary on duration variable
summary(Bank_Fix_Imp$duration)
Bank_Fix_Imp %>%  ggplot(aes(x = duration, fill = term_deposit)) + geom_bar() + ggtitle("Distribution of Duration") + xlab("Duration") +
  scale_x_continuous(breaks = seq(0, 5000, 300)) 
ggplot(Bank_Fix_Imp, aes(x = term_deposit, y=duration, fill=term_deposit)) + geom_boxplot() + ggtitle("Distribution of Duration") + xlab("Response") + ylab ("Duration")
```

## The minimum and maximum values are 0 and 4918 sec and distribution of duration is highly right screwed 
## "duration" and "term_deposit"are pretty strongly associated. The longer duration is, the bigger prportion of people subscibe a term deposit.


```{r }
## Summary on campaign variable. Number of contacts performed during this campaign and for this client 
# summary(Bank_Fix_Imp$campaign)
# Bank_Fix_Imp %>%  ggplot(aes(x = campaign, fill = term_deposit)) + geom_bar() + ggtitle("Distribution of Campaign") + xlab("Campaign")+
#   scale_x_continuous(breaks = seq(0, 50, 1)) 
# ggplot(Bank_Fix_Imp, aes(x = term_deposit, y=campaign, fill=term_deposit)) + geom_boxplot() + ggtitle("Distribution of campaign") + xlab("Response") + ylab ("campaign")
# aggregate(data.frame(count = Bank_Fix_Imp$campaign), list(value = Bank_Fix_Imp$campaign), length)
# Bank_Fix_Imp <- Bank_Fix_Imp %>%   
#     filter(campaign <= 10) 
```

## The minimum and maximum values are 1 and 56 and distribution of campaign is right screwed 
## looks like outlier in capaign varaible, after 8, the outcome is "no" for all observations. we can limit our study to 8
## Most of the campaign is on 1 and 2.
## There is a trend that the more number of campaign, the less percentage of clients substribe a term deposit, Expecially for campaign more than 3.


```{r }
## Summary on pdays variable. Number of days that passed by after the client was last contacted from a previous campaign
summary(Bank_Fix_Imp$pdays)
Bank_Fix_Imp %>%  ggplot(aes(x = pdays, fill = term_deposit)) + geom_bar() + ggtitle("Distribution of pdays") + xlab("pdays") 
ggplot(Bank_Fix_Imp, aes(x = term_deposit, y=pdays, fill=term_deposit)) + geom_boxplot() + ggtitle("Distribution of pdays") + xlab("Response") + ylab ("pdays")
aggregate(data.frame(count = Bank_Fix_Imp$pdays), list(value = Bank_Fix_Imp$pdays), length)
```
## most of the observations has value of 999 which mean these customers never contacted in the past.

```{r }
## Summary on previous variable. How many number of contacts performed before this campaign 
summary(Bank_Fix_Imp$previous)
Bank_Fix_Imp %>%  ggplot(aes(x = previous, fill = term_deposit)) + geom_bar() + ggtitle("Distribution of previous") + xlab("previous") 
ggplot(Bank_Fix_Imp, aes(x = term_deposit, y=previous, fill=term_deposit)) + geom_boxplot() + ggtitle("Distribution of previous") + xlab("Response") + ylab ("previous")
aggregate(data.frame(count = Bank_Fix_Imp$previous), list(value = Bank_Fix_Imp$previous), length)
```
## The minimum and maximum values are 0 and 7. Most of the obserations with 0 value mean the customers never contacted in the past.

```{r }
## Summary on emp.var.rate variable. We can remove this variable from our analysis because of multicolliniarity 
summary(Bank_Fix_Imp$emp.var.rate)
Bank_Fix_Imp %>%  ggplot(aes(x = emp.var.rate, fill = term_deposit)) + geom_bar() + ggtitle("Distribution of emp.var.rate") + xlab("emp.var.rate") 
ggplot(Bank_Fix_Imp, aes(x = term_deposit, y=emp.var.rate, fill=term_deposit)) + geom_boxplot() + ggtitle("Distribution of emp.var.rate") + xlab("Response") + ylab ("emp.var.rate")
```

```{r } 
## Summary on cons.price.idx variable. consumer price index - monthly indicator
summary(Bank_Fix_Imp$cons.price.idx)
Bank_Fix_Imp %>%  ggplot(aes(x = cons.price.idx, fill = term_deposit)) + geom_bar() + ggtitle("Distribution of cons.price.idx") + xlab("cons.price.idx")
ggplot(Bank_Fix_Imp, aes(x = term_deposit, y=cons.price.idx, fill=term_deposit)) + geom_boxplot() + ggtitle("Distribution of cons.price.idx") + xlab("Response") + ylab ("cons.price.idx")
```

## Overall, comsumer price index has some difference in "yes" and "no" among different values
## Minimum and maximum values are 92.20 and 94.77 respectively


```{r }
## Summary on cons.conf.idx variable. consumer confidence index - monthly indicator 
summary(Bank_Fix_Imp$cons.conf.idx)
Bank_Fix_Imp %>%  ggplot(aes(x = cons.conf.idx, fill = term_deposit)) + geom_bar() + ggtitle("Distribution of cons.conf.idx") + xlab("cons.conf.idx")
ggplot(Bank_Fix_Imp, aes(x = term_deposit, y=cons.conf.idx, fill=term_deposit)) + geom_boxplot() + ggtitle("Distribution of cons.conf.idx") + xlab("Response") + ylab ("cons.conf.idx")
```
## Overall, comsumer confidence index has some difference in "yes" and "no" among different values
## Minimum and maximum values are -50.8 and -26.9 respectively

```{r }
## Summary on euribor3m variable. euribor 3 month rate - daily indicator 
summary(Bank_Fix_Imp$euribor3m)
Bank_Fix_Imp %>%  ggplot(aes(x = euribor3m, fill = term_deposit)) + geom_bar() + ggtitle("Distribution of euribor3m") + xlab("euribor3m")
ggplot(Bank_Fix_Imp, aes(x = term_deposit, y=euribor3m, fill=term_deposit)) + geom_boxplot() + ggtitle("Distribution of euribor3m") + xlab("Response") + ylab ("euribor3m")
```

## Minimum and maximum values are 0.634 and 5.045 respectively

```{r }
## Summary on nr.employed variable. We can remove this variable from our analysis because of multicolliniarity 
summary(Bank_Fix_Imp$nr.employed)
Bank_Fix_Imp %>%  ggplot(aes(x = nr.employed, fill = term_deposit)) + geom_bar() + ggtitle("Distribution of nr.employed") + xlab("nr.employed")
ggplot(Bank_Fix_Imp, aes(x = term_deposit, y=nr.employed, fill=term_deposit)) + geom_boxplot() + ggtitle("Distribution of nr.employed") + xlab("Response") + ylab ("nr.employed")







```